---
title: "Introduction to R"
output: learnr::tutorial
runtime: shiny_prerendered
progressive: true
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
#remotes::install_github("rstudio/learnr")
#remotes::install_github("rstudio/gradethis",'force=TRUE')
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Welcome to R!

Welcome to your first `lesson in R`. **What is R you may ask?** R is a language that allows you to conduct statistics and data analysis. R is more of a tool to understand data than a programming language used to build software applications for e.g. like Python. You may wonder why we are teaching you to code and this is simply because it is an employable skill that will be beneficial throughout your degree and in the future (many employers look for this skill in their graduate applicants). 

The aim throughout this module is to teach you skills in R which will allow you to do data science; that is, analysing, visualising and interpreting raw data which is usually from a file or database. We also want you to feel confident using R and enjoy learning something new!

What skills you will learn throughout LIFE223QS module: 

1. How to import data into R
2. How to tidy data 
3. How to transform data
4. How to visualize data using computational tools (called models) in R

But firstly it is important to get to grips with R before we begin any actual statistics. 

## Exercise 1: CALCULATIONS

Lets begin with basic syntax to do mathematical calculations in R. R can do addition, subtraction, multiplication and division with + , - , * and / . Syntax just means the structure of statements in a computer language. 

Lets write your first line of R code to calculate some basic math. Mathematical operations in R follow the standard mathematical order of operations aka BIDMASS. 

*Write the R code required to multiply 5 by 2 and press 'Run Code'. You can then check your answer by pressing 'Submit Answer'. *

If you are stuck, you can press the solution button but have a go before revealing the answer!

```{r five-times-two, exercise=TRUE}
'force=TRUE'

```
```{r five-times-two-solution}
5*2
```
```{r five-times-two-check}
grade_result(
  pass_if(~identical(.result, 10))
)
"Well done!"
```


*Great! Now lets try a slightly trickier equation. Write the R code required to multiply 14 by 5 plus 20 divided by 2. Remember to use the appropriate syntax (+ , - , / and *).*

```{r fourteen-times-five-plus-twenty-divided-two, exercise=TRUE}

```
```{r fourteen-times-five-plus-twenty-divided-two-solution}
14*5+20/2
```
```{r fourteen-times-five-plus-twenty-divided-two-check}
"Good job!"
```

## Exercise 2: COMMENTS

We will now show you how to tell R to ignore parts of your code. Why would we want the computer to ignore a part of their file? Well, comments can: 

1. Help other people reading the code understand it faster

```{r half, exercise=TRUE}
0.5
#The probability of flipping a coin and getting a head
```




2. Give context for why something is written the way it is 

```{r TikTok_count, exercise=TRUE}
TikTok_count < 0
#This code will be used to count the number of times anyone tweets the word TikTok
```



3. Ignore a line of code and see how a program will run without it

```{r}

```


Text written in the program but not run by the computer is called a comment. R interprets anything after a # as a comment.  

*Add a comment inside this code block of how to calculate volume of cube* 

```{r include=TRUE, three-times-three-times-three, exercise=TRUE}
3*3*3

```


## Exercise 3: TYPES OF DATA

In R and programming, data types are the classifications we give different kinds of information pieces. The different types of data we will be working with throughout this module are: 

- Numeric --> Any number with or without a decimal point
- Character --> Any grouping of characters on your keyboard or text. These strings (what they're referred to in R) are surrounded by single quotes '...' or double quotes "..."
- Logical --> This type of data has only two possible values TRUE or FALSE (without the quotes)
- Vectors --> List of related data that is all the same type
- NA --> This type of data represents the absence of a value and is represented by the keyword NA. There can be numeric NA, character NA and a logical NA. 

### Quiz

```{r quiz}
quiz()
  question("What type of data would print('TikTok') belong to?",
           answer("Numeric"),
           answer("Character", correct = TRUE),
           answer("Logical")
)           
```

You may wonder why the word print comes before TikTok. Print() is an example of a function. Functions are "self contained" modules of code that accomplish a specific task. Functions usually take in some sort of data structure (value, vector, dataframe), process it and return a result. In this case, the print function prints its argument and returns it invisibly. 

*Have a go at printing your name as a character string*

```{r print_name_, exercise=TRUE}
print('Imogen')
```

```{r print_name_-hint}
Remember what type of data a name is!
```



## Exercise 4: VARIABLES

Now that we know how R can classify some basic types of information, we will now learn how to store them. Variables allow you to store information and associate that information with a name. In R, we assign variables by using the assignment operator, an arrow sign (<-) created by a carat and a dash. 


*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```



### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```



Some questions to verify that you understand the purposes of various base and recommended R packages:


```

